---
title: "P8105 HW1 - dcm2192"
author: "Dylan Morgan"
date: "2024-09-21"
output: github_document
---

```{r setup, message = FALSE}
library(tidyverse)
library(tidyr)
```

This file will include answers to HW1 questions.

## Problem 1

*Load `penguins` dataset.*
```{r}
data("penguins", package = "palmerpenguins")
```

The `penguins` dataset provides information regarding the locations, classifications, and physical attributes of the penguins. Specifically, it describes the `species`, `island`, `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, `body_mass_g`, `sex`, and `year` of each penguin. The penguins are classified by `species` as either Adelie, Chinstrap, or Gentoo. Each penguin comes from one of three islands from the `island` variable: Biscoe, Dream, and Torgersen. The size of their bills, measured in millimeters, is shown as decimals in `bill_length_mm` and `bill_depth_mm`, respectively. The length of their flippers, measured in millimeters, is shown as integers in `flipper_length_mm`. Their body mass, measured in grams, is shown as integers in `body_mass_g`. Their sex is categorized as either female or male, as shown in `sex`. The `year` column is provided as integers, with values ranging from `r min(pull(penguins, year))` to `r max(pull(penguins, year))`. The `penguins` dataset contains `r nrow(penguins)` observations and `r ncol(penguins)` variables. The mean flipper length is `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)` mm.

*Create scatterplot of flipper length and bill length, with colors categorized by species.*
```{r}
ggplot(penguins, 
       aes(x = bill_length_mm, 
           y = flipper_length_mm, 
           color = species)) + 
  geom_point()
```

*Export scatterplot to project directory.*
```{r}
ggsave("penguins_flipper_bill_plot.pdf")
```

## Problem 2

*Create data frame.*
```{r}
prob2_df <- 
  tibble(
    norm_samp = rnorm(n = 10, mean = 1), 
    samp_greater_than_0 = norm_samp > 0, 
    char_vec = c("John", "Bob", "Mark", "Jesse", "Mary", "Barbie", "Molly", "Terry", "Tom", "Ralph"), 
    factor_vec = factor(c("urban", "urban", "rural", "suburban", "suburban", "urban", "rural", "urban", "rural", "urban"))
  )

prob2_df
```

*Try to take the mean of each variable.*
```{r}
mean(pull(prob2_df, norm_samp))
mean(pull(prob2_df, samp_greater_than_0))
mean(pull(prob2_df, char_vec))
mean(pull(prob2_df, factor_vec))
```

In the `prob2_df` data frame, the numeric columns `norm_samp` and `samp_greater_than_0` are able to generate a mean, whereas the character vector `vector` and the factor vector `factor_vec` are unable to generate a mean. This is because R cannot calculate the average value of non-numeric variables. 

*Try to convert each variable to numeric.*
```{r, eval = FALSE}
as.numeric(pull(prob2_df, n orm_samp))
as.numeric(pull(prob2_df, samp_greater_than_0))
as.numeric(pull(prob2_df, char_vec))
as.numeric(pull(prob2_df, factor_vec))
```

The `as.numeric` function converts a non-numeric variable into a numeric variable, and it is successful in doing so for the variables `norm_samp`, `samp_greater_than_0`, and `factor_vec` in the `prob2_df` data frame. This is because `norm_samp` is already a numeric variable, `samp_greater_than_0` is a logical variable that can be converted to a binary true or false (1 or 0) value, and `factor_vec` is a factor variable whose category can be assigned a numerical value based on the number of levels (in this case, 3). `as.numeric` cannot be applied to `char_vec` because this variable is non-numeric without any possible categorization, since each string value in this column has a unique value and cannot be reclassified like the logical and factor variables can. This also explains why R cannot generate the mean of `sample_greater_than_0`, `char_vec`, and `factor_vec` without first converting them to numeric values, as doing so would allow R to use these numbers to perform the average calculation for each variable. 
