---
title: "P8105 HW1 - dcm2192"
author: "Dylan Morgan"
date: "2024-09-21"
output: github_document
---

```{r setup, message = FALSE}
library(tidyverse)
library(tidyr)
```

This file will include answers to HW1 questions.

## Problem 1

*Load `penguins` dataset.*
```{r}
data("penguins", package = "palmerpenguins")
```

The `penguins` dataset provides information regarding the locations, classifications, and physical attributes of the penguins. Specifically, it describes the `species`, `island`, `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, `body_mass_g`, `sex`, and `year` of each penguin. The penguins are classified by `species` as either Adelie, Chinstrap, or Gentoo. Each penguin comes from one of three islands from the `island` variable: Biscoe, Dream, and Torgersen. The size of their bills, measured in millimeters, is shown as decimals in `bill_length_mm` and `bill_depth_mm`, respectively. The length of their flippers, measured in millimeters, is shown as integers in `flipper_length_mm`. Their body mass, measured in grams, is shown as integers in `body_mass_g`. Their sex is categorized as either female or male, as shown in `sex`. The `year` column is provided as integers, with values ranging from `r min(pull(penguins, year))` to `r max(pull(penguins, year))`. The `penguins` dataset contains `r nrow(penguins)` observations and `r ncol(penguins)` variables. The mean flipper length is `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)` mm.

*Create scatterplot of flipper length and bill length, with colors categorized by species.*
```{r}
ggplot(penguins, 
       aes(x = bill_length_mm, 
           y = flipper_length_mm, 
           color = species)) + 
  geom_point()
```

*Export scatterplot to project directory.*
```{r}
ggsave("penguins_flipper_bill_plot.pdf")
```

## Problem 2

*Create data frame.*
```{r}
prob2_df <- 
  tibble(
    norm_samp = rnorm(n = 10, mean = 1), 
    samp_greater_than_0 = norm_samp > 0, 
    char_vec = c("John", "Bob", "Mark", "Jesse", "Mary", "Barbie", "Molly", "Terry", "Tom", "Ralph"), 
    factor_vec = factor(c("urban", "urban", "rural", "suburban", "suburban", "urban", "rural", "urban", "rural", "urban"))
  )
```



